The Noisy Work of Uncertainty Visualisation Research: A Review

### 2024-04-27 paper.qmd
added bivariate pallette map (kind of) and attempted VSUP map

```{r}

```

>>>>

```{r}
#| eval: false
#| echo: false

# seed for sampling
set.seed(1997)

# load libraries
library(strayr)
library(tidyverse)
library(biscale)
library(RColorBrewer)

# Get map data
#sa4 <- read_absmap("sa42021")
state <- read_absmap("state2021") |>
  na.omit() |>
  dplyr::mutate(temp = 31 - 0.001*(cent_long-110)^(2) - 0.003*(cent_lat-12.52294)^(2), # trend
         notrend = rnorm(1), # no trend
         highvar = runif(9, min=2, max=4), # high variance
         lowvar = runif(9, min=0, max=2), # low variance
         ) |>
  pivot_longer(cols=highvar:lowvar, names_to = "variance_class", values_to = "variance") |>
  # add bivariate classes to data
  mutate(bitemp = cut(temp, breaks=20:28, labels=seq(8)),
         bivar = cut(variance, breaks=0:4, labels=seq(4)),
         biclass = paste(bitemp, bivar, sep="-")
  )

#sa4 <- sa4 |>
#  dplyr::mutate(temp = 30 - 0.001*(cent_long-123.7799)^(2) - 0.003*(cent_lat-12.52294)^(2), # trend
#         notrend = rnorm(1), # no trend
#         highvar = runif(1, min=4, max=5), # high variance
#         lowvar = runif(1, min=0, max=1), # low variance
#         )

# Choropleth Map
state %>%
  ggplot() +
  geom_sf(aes(fill = temp, geometry = geometry)) + 
  scale_fill_distiller(type = "seq", palette = "YlOrRd", direction = 1)+
  facet_wrap(vars(variance_class))

# Bivariate Map
# Make bivariate palette
# Function to devalue by a certain amount
colsupress <- function(basecols, hue=1, sat=1, val=1) {
    X <- diag(c(hue, sat, val)) %*% rgb2hsv(col2rgb(basecols))
    hsv(pmin(X[1,], 1), pmin(X[2,], 1), pmin(X[3,], 1))
}

# set Base colours
basecols <- brewer.pal(8, name = "YlOrRd")
# recurvisely decrease value
v_val = 0.5
bivariatepal <- c(basecols,
                  colsupress(basecols, sat=v_val),
                  colsupress(colsupress(basecols, sat=v_val), sat=v_val),
                  colsupress(colsupress(colsupress(basecols, sat=v_val), sat=v_val), sat=v_val))
# establish levels of palette
names(bivariatepal) <- paste(rep(1:8, 4), "-" , rep(1:4, each=8), sep="")
# preview palette
show_col(bivariatepal, ncol=8)

# Bivariate maps
state |>
  ggplot() +
  geom_sf(aes(fill = biclass, geometry = geometry)) + 
  scale_fill_manual(values = bivariatepal) +
  facet_wrap(vars(variance_class))

# VSUP
# Function to combine hue values
evenhueblend <- function(basecols, p_length, nblend) {
    X <- rgb2hsv(col2rgb(unique(basecols)))
    v1 <- X[,seq(1,dim(X)[2], 2)]
    v2 <- X[,seq(2,dim(X)[2], 2)]
    v3 <- (v1 + v2)
    if("matrix" %in% class(v3)){
      v3["h",] <- ifelse(v3["h",]>=1, v3["h",]-1 ,v3["h",])/2 # hue is on a circle so 0 wraps around to 1, need to do differently
      v3 <- v3/2
      nsup <- nblend
      hsv(rep(v3[1,], each=nsup), rep(v3[2,], each=nsup), rep(v3[3,], each=nsup))
      } else {
        rep(hsv(h=v3[1], s=v3[2], v=v3[3]), p_length)
        }
}
VSUPfunc <- function(basecols, p_length, nblend){
  evenhueblend(colsupress(basecols, sat=0.5), p_length, nblend)
}
# VSUP
p = length(basecols)
VSUP <- c(basecols,
          VSUPfunc(basecols, p, 2),
          VSUPfunc(VSUPfunc(basecols, p, 2), p, 4),
          VSUPfunc(VSUPfunc(VSUPfunc(basecols, p, 2), p, 4), p, 8))
names(VSUP) <- paste(rep(1:8, 4), "-" , rep(1:4, each=8), sep="")
show_col(VSUP, ncol=8)
# Pixel Map
# Sample Map
```
